\section*{NeurIPS Paper Checklist}

\begin{enumerate}

\item {\bf Claims}
    \item[] Question: Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \item[] Answer: \answerYes{}
    \item[] Justification: The abstract and introduction summarize the matched-accuracy discriminator result, the theory contribution, the human-validation study, and the scope limits on model coverage. Limitations and scope restrictions are also stated explicitly in the introduction and the Limitations section.

\item {\bf Limitations}
    \item[] Question: Does the paper discuss the limitations of the work performed by the authors?
    \item[] Answer: \answerYes{}
    \item[] Justification: Section~\ref{sec:limitations} discusses judge overcalling, the load-bearing nature of the 9-point scale, model-coverage gaps, limited human-sample size, failed predictive calibration, benchmark misuse, and the fact that queries are LLM-generated.

\item {\bf Theory assumptions and proofs}
    \item[] Question: For each theoretical result, does the paper provide the full set of assumptions and a complete (and correct) proof?
    \item[] Answer: \answerYes{}
    \item[] Justification: Theorem~\ref{thm:nonreduce} and Proposition~\ref{prop:resolution} are stated in Section~\ref{sec:theory}, and complete proofs are provided in Appendix~\ref{app:proofs}. The main text includes proof sketches for intuition and the appendix provides the formal derivations.

\item {\bf Experimental result reproducibility}
    \item[] Question: Does the paper fully disclose all the information needed to reproduce the main experimental results of the paper to the extent that it affects the main claims and/or conclusions of the paper (regardless of whether the code and data are provided or not)?
    \item[] Answer: \answerYes{}
    \item[] Justification: Sections~\ref{sec:method} and \ref{sec:setup}, together with the appendices and supplemental artifact, describe the benchmark construction, model catalog, scoring pipeline, fitting procedure, and statistical analyses used for the main claims. The supplemental artifact also includes scripts and a reproduction guide for rerunning the analyses from saved data.

\item {\bf Open access to data and code}
    \item[] Question: Does the paper provide open access to the data and code, with sufficient instructions to faithfully reproduce the main experimental results, as described in supplemental material?
    \item[] Answer: \answerYes{}
    \item[] Justification: The anonymized supplemental artifact includes the code, saved analysis outputs, released data files, Croissant metadata, and reproduction instructions. The paper states that the benchmark, scoring toolkit, human-validation data, and analysis code are released for replication.

\item {\bf Experimental setting/details}
    \item[] Question: Does the paper specify all the training and test details (e.g., data splits, hyperparameters, how they were chosen, type of optimizer) necessary to understand the results?
    \item[] Answer: \answerYes{}
    \item[] Justification: The paper specifies the 10,000-query benchmark design, domain/tier stratification, model catalog, decoding setup, severity scale, dual-judge resolution pipeline, human-validation protocol, candidate distribution families, and bootstrap procedures. Additional tables and prompts are included in the appendix and supplemental files.

\item {\bf Experiment statistical significance}
    \item[] Question: Does the paper report error bars suitably and correctly defined or other appropriate information about the statistical significance of the experiments?
    \item[] Answer: \answerYes{}
    \item[] Justification: The paper reports bootstrap confidence intervals for \sdi{}, p-values for correlation and goodness-of-fit tests, mutual-information decomposition, and additional robustness analyses. The bootstrap setup and significance thresholds are described in the paper and supplemental scripts.

\item {\bf Experiments compute resources}
    \item[] Question: For each experiment, does the paper provide sufficient information on the computer resources (type of compute workers, memory, time of execution) needed to reproduce the experiments?
    \item[] Answer: \answerNo{}
    \item[] Justification: The paper states that inference was run through an open-access third-party API hosting the target models, but it does not provide exact GPU types, memory, or wall-clock runtime for every experiment. We chose not to over-claim compute reproducibility where the provider abstraction hides the underlying hardware.

\item {\bf Code of ethics}
    \item[] Question: Does the research conducted in the paper conform, in every respect, with the NeurIPS Code of Ethics \url{https://neurips.cc/public/EthicsGuidelines}?
    \item[] Answer: \answerYes{}
    \item[] Justification: The submission is anonymized, the paper discusses misuse risks and limitations, and the released artifact is intended for auditing and evaluation rather than unsafe deployment. We are not aware of any aspect of the work that conflicts with the NeurIPS Code of Ethics.

\item {\bf Broader impacts}
    \item[] Question: Does the paper discuss both potential positive societal impacts and negative societal impacts of the work performed?
    \item[] Answer: \answerYes{}
    \item[] Justification: Section~\ref{sec:limitations} includes a Broader impacts paragraph discussing positive impacts for risk auditing and deployment gating, as well as negative impacts from benchmark gaming or optimizing for better-looking but still dangerous failure profiles.

\item {\bf Safeguards}
    \item[] Question: Does the paper describe safeguards that have been put in place for responsible release of data or models that have a high risk for misuse (e.g., pre-trained language models, image generators, or scraped datasets)?
    \item[] Answer: \answerNA{}
    \item[] Justification: The paper releases an evaluation benchmark, scores, and analysis code, not a new generative model or scraped multimedia dataset with elevated release risk. We therefore view the high-risk-model safeguard question as not directly applicable to this artifact.

\item {\bf Licenses for existing assets}
    \item[] Question: Are the creators or original owners of assets (e.g., code, data, models), used in the paper, properly credited and are the license and terms of use explicitly mentioned and properly respected?
    \item[] Answer: \answerNo{}
    \item[] Justification: The paper cites the upstream benchmarks and model families used in the evaluation, and the supplemental artifact includes licenses for the released ERRORQUAKE code/data. However, the submission does not enumerate every upstream model or provider license directly in the manuscript, so we answer conservatively.

\item {\bf New assets}
    \item[] Question: Are new assets introduced in the paper well documented and is the documentation provided alongside the assets?
    \item[] Answer: \answerYes{}
    \item[] Justification: The benchmark release includes a datasheet, Croissant metadata, reproduction instructions, and scripts for the reported analyses. The paper and supplemental material document the query structure, severity scale, scoring fields, and released outputs.

\item {\bf Crowdsourcing and research with human subjects}
    \item[] Question: For crowdsourcing experiments and research with human subjects, does the paper include the full text of instructions given to participants and screenshots, if applicable, as well as details about compensation (if any)? 
    \item[] Answer: \answerNo{}
    \item[] Justification: The supplemental artifact includes the full rating instructions used in the human-validation study. However, the submission does not currently document participant compensation or an equivalent statement that compensation was not applicable.

\item {\bf Institutional review board (IRB) approvals or equivalent for research with human subjects}
    \item[] Question: Does the paper describe potential risks incurred by study participants, whether such risks were disclosed to the subjects, and whether Institutional Review Board (IRB) approvals (or an equivalent approval/review based on the requirements of your country or institution) were obtained?
    \item[] Answer: \answerNo{}
    \item[] Justification: The submission reports a human-rating study but does not include an IRB-approval statement or a fuller discussion of participant-risk disclosure. We therefore answer conservatively rather than implying review documentation that is not present in the anonymized materials.

\item {\bf Declaration of LLM usage}
    \item[] Question: Does the paper describe the usage of LLMs if it is an important, original, or non-standard component of the core methods in this research? Note that if the LLM is used only for writing, editing, or formatting purposes and does \emph{not} impact the core methodology, scientific rigor, or originality of the research, declaration is not required.
    \item[] Answer: \answerYes{}
    \item[] Justification: The paper explicitly states that the benchmark queries were generated by a frontier LLM and that response scoring uses an LLM-judge pipeline with human validation. These components are described in Sections~\ref{sec:method}, \ref{sec:setup}, and \ref{sec:limitations}.

\end{enumerate}
